Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-07 vesicle_code/docs/adhesion/adhesion.tex
edited up to end of section II

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral Equation Formulation}
Because the fluid equations are linear and elliptic, a boundary integral
equation formulation is possible.  This has several advantages,
including that only the vesicle interfaces require discretization.  We
start by defining the Stokes single-layer potential
\begin{align}
  \SS[\ff](\xx) &= \frac{1}{4\pi\mu} \int_\gamma \left(
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff(\yy) ds_\yy, \quad \xx \in \RR^2,
  \label{eqn:SLP}
\end{align}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$.  Then, given a vesicle
with a traction jump $\xxi$, the fluid velocity at is~\cite{poz1992}
\begin{align*}
  \uu(\xx) = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \quad \xx \in \RR^2.
\end{align*}
The traction jump is
\begin{align*}
  \xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA \xx,
\end{align*}
where the three terms correspond to the forces due to bending, tension,
and adhesion, respectively.  Therefore Applying the no-slip boundary
condition on each vesicle, the governing equations are
\begin{align*}
  &\dot{\xx} = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \\
  &\xx_s \cdot \uu_s = 0, \\
  &\xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA\xx.
\end{align*}
Defining the bending operator as $\BB[\xx](\ff) = -\kappa_b \ff_{ssss}$,
and the tension operator $\TT[\xx](\sigma) = (\sigma \xx_s)_s$, the
no-slip boundary condition gives
\begin{align*}
  \pderiv{\xx}{t} = \SS \BB \xx + \SS \TT \sigma + \SS \AA \xx.
\end{align*}
We apply the IMEX-Euler method 
\begin{align*}
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \SS^N \BB^N \xx^{N+1} + 
  \SS^N \TT^N \sigma^{N+1} + \SS^N \AA^N \xx^N,
\end{align*}
that treats both the bending and tension terms
semi-implicitly~\cite{qua-bir2014}. Note that the adhesive term is
treated explicitly.  Therefore, a single time step requires solving
\begin{align*}
  \xx^{N+1} - \Delta t \SS^N \BB^N \xx^{N+1} - 
    \Delta t \SS^N \TT^N \sigma^{N+1} = \xx^N + 
    \Delta t \SS^N \AA^N \xx^N,
\end{align*}
along with the inextensibility constraint that is discretized as
\begin{align*}
  \xx_s^{N} \cdot \xx_{s}^{N+1} = 1.
\end{align*}
We discretize the vesicles at a set of collocation points, compute the
bending and tension terms with Fourier differentiation, and apply Alpert
quadrature~\cite{alp1999} to the logarithmically singular single-layer
potential $\SS$.  The source and target points of the adhesion force
never coincide since they are always on different vesicles, so the
adhesion force~\eqref{eqn:adhesionForce} is computed with the spectrally
accurate trapezoid rule~\cite{tre-wei2014}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Shear and Normal Stresses}
To characterize the rheological properties of the suspension of a
doublet, it is necessary to compute shear and normal stresses.  At a
point $\xx$ in the fluid bulk, The velocity $\uu(\xx)$, pressure
$p(\xx)$, and total stress tensor $T = - p I + \nabla \uu + \nabla
\uu^T$ of the single-layer potential~\eqref{eqn:SLP} are
\begin{align*}
  \uu(\xx) &= \SS_\uu[\ff](\xx) = \frac{1}{4\pi\mu}\int_{\gamma} \left( 
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff ds_\yy, \\
    p(\xx) &= \SS_p[\ff](\xx) = \frac{1}{2\pi} \int_{\gamma} 
    \frac{\rr \cdot \ff}{\rho^2} ds_\yy, \\
    T(\xx) &= \SS_T[\ff](\xx) = -\frac{1}{\pi}\int_{\gamma}
      \frac{\rr \cdot \ff}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2} ds_\yy.
\end{align*}
To compute the stress tensor on a vesicle, there is a jump in the layer
potential $\SS_T$, and the stress tensor is
\begin{align*}
  T(\xx) = \frac{1}{2}(\nn \otimes \ff) + \frac{1}{2} \left(
    \tt \otimes \left[
    \begin{array}{cc}
      2 t_x t_y & t_y^2 - t_x^2 \\
      t_y^2 - t_x^2 & -2 t_x t_y
    \end{array}
    \right] \ff \right) + 
  \SS_T[\ff](\xx), \quad \xx \in \gamma,
\end{align*}
where $\nn$ and $\tt$ are the vesicle's outward unit normal and tangent
vector, respectively.  With the stress tensor computed for $\xx \in
\gamma$, the shear and normal stresses are
\begin{align*}
  \tau^\tt = (T \nn) \cdot \tt, \qquad
  \tau^\nn = (T \nn) \cdot \nn.
\end{align*}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral Equation Formulation}
Because the fluid equations are linear and elliptic, a boundary integral
equation formulation is possible.  This has several advantages,
including that only the vesicle interfaces require discretization.  We
start by defining the Stokes single-layer potential
\begin{align}
  \SS[\ff](\xx) &= \frac{1}{4\pi\mu} \int_\gamma \left(
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff(\yy) ds_\yy, \quad \xx \in \RR^2,
  \label{eqn:SLP}
\end{align}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$.  Then, given a vesicle
with a traction jump $\xxi$, the fluid velocity is~\cite{poz1992}
\begin{align*}
  \uu(\xx) = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \quad \xx \in \RR^2.
\end{align*}
The traction jump is
\begin{align*}
  \xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA \xx,
\end{align*}
where the three terms correspond to the forces due to bending, tension,
and adhesion, respectively.  Therefore, applying the no-slip boundary
condition on each vesicle, the governing equations are
\begin{align*}
  &\dot{\xx} = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \\
  &\xx_s \cdot \uu_s = 0, \\
  &\xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA\xx.
\end{align*}
Defining the bending operator as $\BB[\xx](\ff) = -\kappa_b \ff_{ssss}$,
and the tension operator $\TT[\xx](\sigma) = (\sigma \xx_s)_s$, the
no-slip boundary condition gives
\begin{align*}
  \pderiv{\xx}{t} = \SS \BB \xx + \SS \TT \sigma + \SS \AA \xx.
\end{align*}
We apply the IMEX-Euler method 
\begin{align*}
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \SS^N \BB^N \xx^{N+1} + 
  \SS^N \TT^N \sigma^{N+1} + \SS^N \AA^N \xx^N,
\end{align*}
that discretizes both the bending and tension terms
semi-implicitly~\cite{qua-bir2014}, and the adhesive term explicitly.
Therefore, a single time step requires solving
\begin{align*}
  \xx^{N+1} - \Delta t \SS^N \BB^N \xx^{N+1} - 
    \Delta t \SS^N \TT^N \sigma^{N+1} = \xx^N + 
    \Delta t \SS^N \AA^N \xx^N,
\end{align*}
along with the inextensibility constraint that is discretized as
\begin{align*}
  \xx_s^{N} \cdot \xx_{s}^{N+1} = 1.
\end{align*}
We discretize the vesicles at a set of collocation points, compute the
bending and tension terms with Fourier differentiation, and apply Alpert
quadrature~\cite{alp1999} to the weakly-singular single-layer potential
$\SS$.  The source and target points of the adhesion force never
coincide since they are always on different vesicles, so the adhesion
force~\eqref{eqn:adhesionForce} is computed with the spectrally accurate
trapezoid rule~\cite{tre-wei2014}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Shear and Normal Stresses}
To characterize the rheological properties of the suspension of a
doublet, it is necessary to compute the shear and normal stresses.  At a
point $\xx \in \RR^2 \backslash \gamma$ in the fluid bulk, the velocity
$\uu(\xx)$, pressure $p(\xx)$, and total stress tensor $T = - p I +
\nabla \uu + \nabla \uu^T$ of the single-layer potential~\eqref{eqn:SLP}
are
\begin{align*}
  \uu(\xx) &= \SS_\uu[\ff](\xx) = \frac{1}{4\pi\mu}\int_{\gamma} \left( 
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff ds_\yy, \\
    p(\xx) &= \SS_p[\ff](\xx) = \frac{1}{2\pi} \int_{\gamma} 
    \frac{\rr \cdot \ff}{\rho^2} ds_\yy, \\
    T(\xx) &= \SS_T[\ff](\xx) = -\frac{1}{\pi}\int_{\gamma}
      \frac{\rr \cdot \ff}{\rho^2} 
      \frac{\rr \otimes \rr}{\rho^2} ds_\yy.
\end{align*}
To compute the stress tensor on a vesicle, there is a jump in the layer
potential $\SS_T$, and the stress tensor is
\begin{align*}
  T(\xx) = \frac{1}{2}(\nn \otimes \ff) + \frac{1}{2} \left(
    \tt \otimes \left[
    \begin{array}{cc}
      2 t_x t_y & t_y^2 - t_x^2 \\
      t_y^2 - t_x^2 & -2 t_x t_y
    \end{array}
    \right] \ff \right) + 
  \SS_T[\ff](\xx), \quad \xx \in \gamma,
\end{align*}
where $\nn$ and $\tt$ are the vesicle's outward unit normal and tangent
vector, respectively.  With the stress tensor computed for $\xx \in
\gamma$, the shear and normal stresses are
\begin{align*}
  \tau^\tt = (T \nn) \cdot \tt, \qquad
  \tau^\nn = (T \nn) \cdot \nn.
\end{align*}
